# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo = TRUE}
library(plyr)
library(lattice)
 #reads in the file and subsets to the requested dates
  data <- read.csv("activity.csv")
  #Data Processing Steps                
  
```

## What is mean total number of steps taken per day?
```{r echo = TRUE}
 ##What is mean total number of steps taken
  #Ignore missing values
  datasub <- !is.na(data$steps)
  data2 <- data[datasub,]

#Make a histogram of total number of steps taken each day
#Calculating the average steps by day  
d <- ddply(data2, .(date), summarize, walk = sum(steps))
  
  hist(d$walk, col = "red", 
       main = "Histogram of Total Steps by Day",  
       xlab = "Total Number of Steps",
       ylab = "Frequency (# days)",
       breaks = 15)

  #Calculate mean and median
  removedmean <- mean(d$walk)
  removedmedian <- median(d$walk)
```

The mean for the data not including the missing values is `r removedmean`. 

The median for the data not including the missing values is `r removedmedian`.

## What is the average daily activity pattern?
```{r echo = TRUE}
 ##What is the average daily activity pattern
  #Time series plot of 5 minute interval and average number of steps
  e <- ddply(data2, .(interval), summarize, steps = mean(steps))
  plot(e, type = "l",
       main = "Time Series Plot of Average Steps",  
       xlab = "Interval (min)",
       ylab = "Average (steps)")
  #Which 5 minute interval, on average accross all days in the data 
  #set contains the maximum number of steps?
  emaxinterval <- e[which.max(e[,2]),1]
```
Interval `r emaxinterval` has the maximum number of steps averaged accross all days.

## Imputing missing values
```{r echo = TRUE}
 ##Imputing missing values
  #Calculate and report the total number of missing values in dataset
  missingvalues <- sum(!datasub)
```
There are `r missingvalues` missing values in the original dataset.

```{r echo = TRUE}
#Devise a stratagy for filling in all of the missing values in the data set.
missingtable <- table(data2$date)
```
The data is missing in full days. 

Method of replacing the missing data:
NA values replaced with steps averaged over intervals matching intervals.

```{r echo = TRUE}

#Create a new dataset that is equal to the original dataset but with the missing data filled in.
#Using a method of replacing the missing data from the average daily steps.
data$date <- strptime(data$date[], format = "%Y-%m-%d")  
data3 <- data

  for (i in 1:length(data3$steps)){
    if(is.na(data3$steps[i])) 
      data3$steps[i] <- e$steps[which(e$interval==data3$interval[i])]
    }

  d2 <- ddply(data3, .(date), summarize, walk = sum(steps))
  hist(d2$walk, col = "blue", 
       main = "Histogram of Total Steps by Day (replaced data)",  
       xlab = "Total Number of Steps",
       ylab = "Frequency (# days)",,
       breaks = 15)
  
  #Calculate mean and median
  newmean <- mean(d2$walk)
  newmedian <- median(d2$walk)
```
The mean for the data with replaced values is `r newmean`.

The median for the data with replaced values is `r newmedian`.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
 weekpart <- (weekdays(data3$date) == "Sunday" | weekdays(data3$date) == "Saturday")
  data3$weekpart[weekpart] <- "weekend"
  data3$weekpart[!weekpart] <- "weekday"
  
  f <- ddply(data3, .(interval, weekpart), summarize, steps = mean(steps))

  xyplot(f$steps~f$interval|f$weekpart, type = "l", 
         layout = c(1,2), 
         main = "Weekend versus Weekday, Time Series of Steps by Interval",  
         xlab = "Interval",
         ylab = "Number of Steps")
```

 
  
 
  
 
  
 
 
  
  
 